---
title: "Data frames"
author: "Ines"
date: '2022/23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  Clearing the environment and library loading

Clearing the environment.

```{r}
rm(list=ls())
```

Loading library tidyverse.

```{r}
library(tidyverse)
```

## Creating a data frame from scratch

```{r}
caseID <- c("C1", "C2", "C3", "C4")
problem <- c("theft", "robbery", "assault", "robbery")
solved <- c(TRUE, TRUE, FALSE, TRUE)
courtDays <- c(3,2,1,14)
courtCases <- 
	data.frame(caseID, problem, solved, courtDays, stringsAsFactors=T)
```

Checking the result

```{r}

View(courtCases)
```


```{r}
courtCases$problem
```

## Creating a data frame from 2 data frames

We created a data frame courtCases earlier. Here, we create a 2nd one with the same column names (casesID, problem, solved, courtDays). This time we have 3 new instances (records).

```{r}
caseID <- c("C5", "C6", "C7")
problem <- c("robbery","theft", "robbery")
solved <- c(TRUE, FALSE, TRUE)
courtDays <- c(5,1,2)
newCourtCases <- 
	data.frame(caseID, problem, solved, courtDays, stringsAsFactors=T)
```



## Adding data frames courtCases and newCourtCases together


The courtCases and newCourtCases data frames can be put together by binding rows with rbind()

```{r}
allCourtCases <- rbind(courtCases, newCourtCases)
```

You can check that the binding has occurred using View() or the name of the new data frame.

```{r}
View(allCourtCases)
```

## Adding columns

assume you want to add the waiting time before the case went to court.

```{r}
wait <- c(45,53,80,101,49,77,110)
```


To add this to the allCourtCases dataframe

```{r}
allCourtCases$wait <- wait
```

You can now add 


```{r} 
allCourtCases$time <- allCourtCases$wait + allCourtCases$courtDays
```


```{r} 
allCourtCases$threePlus <- allCourtCases$time > 90 
```

## Creating a data frame from a file

### Loading data file onto a dara frame

The code below reads data into a data frame. 


```{r}
labor <- read.csv("labor.csv", header = FALSE, stringsAsFactors=T)
```

When you inspect this dataset you will see that there are a number of NA values. This means there is no value for that specific attribute (column) and instance (row).


### Loading a data file onto a tibble

```{r}
laborT <- read_csv("labor.csv" )
laborT <- laborT  |> mutate(across(where(is.character),as_factor))
```


### Converting a tibble to a data frame

```{r}
class(laborT)
laborTtoDF <- as.data.frame(laborT)
class(laborTtoDF)
```

### Converting a data frame to a tibble

```{r}

class(labor)
laborToT <- as_tibble(labor)
class(laborToT)

```



## Using R’s pre-installed datasets

Checking the list of datasets included in R

```{r}
data() 
```

One existing dataset is BOD. To load the BOD (biochemical oxygen demand)  dataset  use

```{r}
data(BOD)
```

To see what this dataset contains 

```{r}
View(BOD)
```
To check its type, i.e. that it is a data.frame

```{r}
class(BOD)
```

Another dataset is the co2 dataset. To inspect the first 3 rows of the co2 dataset

```{r}
head(CO2, n = 3) 
```

To inspect the last 3 rows of the co2 dataset

```{r}
tail(CO2, n = 3) 
```

To check the number of columns of the co2 dataset

```{r}
ncol(CO2) 
```

To check the number of rows of the co2 dataset

```{r}
nrow(CO2) 
```

To check both rows and columns use

```{r}
dim(CO2)
```

## Extracting information from a data frame

To get the column names

```{r}
names(CO2) 
```

To get column and row names

```{r}
attributes(CO2) 
```

To get a good idea of the characteristics of a dataset use

```{r}
summary(CO2)
```

To get all the values for the Treatment column in the CO2 dataset

```{r}
CO2$Treatment
```

To get a dataset like CO2 but with only the 4th (“conc”) column

```{r}
CO2fourth <- CO2[,4]
```

To get a dataset like CO2 but with only the 3rd and 4th columns

```{r}
CO2thirdFourth <- CO2[,c(3,4)]
```

To get a dataset like CO2 but with only rows 10 to 20

```{r}
CO2Rows10to20 <- CO2[c(10:20),]
```

Combining both


```{r}
CO2extract <- CO2[c(10:20),c(3,4)]
```

## Excluding specific rows and columns - building a dataset from specific columns and rows

To get a dataset like CO2 but with no Treatment column (the 3rd one) 

```{r}
CO2NoTreat <- CO2[,-3]
```

To get a dataset like CO2 but excluding rows 4 to 9

```{r}
CO2NoRows4to9 <- CO2[-c(4:9),]
```

Combining both

```{r}
CO2NoTreatNoRows4to9 <- CO2[-c(4:9),3]
```

## Instance (observation) selection

Assuming that we want instances where the Treatment is “chilled”.

```{r}
CO2chilled <- subset(CO2, Treatment =="chilled")
```

Column Treatment is useless (all values are the same) so can be deleted.

```{r}
CO2chilled$Treatment <- NULL
```

We want to use CO2 data, but only where the treatment is chilled and the uptake is at least 30.

```{r}
CO2selectedVals <- subset(CO2, Treatment =="chilled" & uptake >=30)
```

##Column deletion

Column Treatment is useless (all values are the same) so can be deleted.

```{r}
CO2selectedVals$Treatment <- NULL
```

## Exercises with data frames


### Exercise 1

First  a vector for each of the 5 attribute is created (course, module1, credits1, module2 and credits2), with a value for each of the 4 instances. Then the 5 vectors are put together into a dataframe. 

```{r}
course <- c("Data Science", "Cyber Security", "IT", "Digital Media")
module1 <- c("analytics", "ethical hacking", "databases", "3D animation")
credits1 <- c(15,15,20,30)
module2 <- c("project", "forensics","Java programming", "perception")
credits2 <- c(45,15,20,12)
courses <- data.frame(course, module1, credits1,module2, credits2, stringsAsFactors = T)
# to check
View(courses)

```


### Exercise 2

The new instance is created

```{r}
e <- data.frame("Networks", "CISCO CCNA", 25, "Operating systems",40, stringsAsFactors=T)
# Give same column names as the courses data frame
names(e) <- names(courses)
# add the rows in the new data frame to the courses data frame 
courses <- rbind(courses, e)
# to check
View(courses)
```


### Exercise 3

```{r}
courses$total <- courses$credits1 + courses$credits2
# to check
View(courses)
```

### Exercise 4

```{r}
courses$pgcert <- courses$total >=60
# to check
View(courses)
```

### Exercise 5

```{r}
twentyPlus <- subset(courses, courses$credits1 >=20 | courses$credits2 >=20)
# to check
View(twentyPlus)
```

### Exercise 6

```{r}
noDBs <- subset(courses, courses$module1 != "databases" & courses$module2 != "databases")
# to check
View(noDBs)
```


### Exercise 7

```{r}
lenses <- read.csv("contactLenses.csv", stringsAsFactors = T)
```

### Exercise 8

```{r}
head(lenses, 4)

```

### Exercise 9

```{r}
tail(lenses,3)

```

### Exercise 10

```{r}
# to view a summary (data types can be inferred from this)
summary(lenses)

## to check the structure/type
str(lenses)

```
It can be seen that there are no missing values.

### Exercise 11

```{r}
# check value before change -value is myope
lenses[3,]
# change value to pre-presbyopic
lenses$age[3] <- "pre-presbyopic"
# alternative ensuring string is a factor
# lenses$age[3] <- as.factor("pre-presbyopic")
# checking result
lenses[3,]

```

### Exercise 12

```{r}
lenses$tearProductionRate <- NULL
```

### Exercise 13

There are 2 alternatives below. The second one is commented out.You will learn a 3rd one in the next lab.

```{r}
# Option 1
colnames(lenses)[colnames(lenses)=="contactLenses"] <- "recommendation"

# Option 2 
# Add new column
#lenses$recommendation <-  lenses$contactLenses
# Remove old column
#lenses$contactLenses <- NULL

# To check it works
names(lenses)
```


### Exercise 14

This solution uses a ``provisional" column  called astigTrue to hold intermediate results.

```{r}
lenses$astigTrue[lenses$astigmatism =="yes"] <- 1
lenses$astigTrue[lenses$astigmatism =="no"] <- 0
#copy  provisional column
lenses$astigmatism <- lenses$astigTrue
#delete provisional column
lenses$astigTrue <- NULL
# To check it works
head(lenses,6)
```

### Exercise 15

The code is identical to the code you have used earlier with data frames, but using tibbles instead.