---
title: "Lab9&10_Apriori"
author: "Dulla"
date: "2024-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(arules)
library(arulesViz)
```

```{r}
data("Groceries")
summary(Groceries)
```

```{r}
rules <- apriori(Groceries, parameter=list(support=0.001, confidence=0.5))
```

```{r}
inspect(head(sort(rules, by ="lift"),10))
```

```{r}
inspect(head(sort(rules, by ="confidence"),15))
```

```{r}
quality(rules)
```

```{r}
head(quality(rules))
```

###Removing redundant rules
Some rules may be redundant (covered by other rules). You may remove them, using a quality measure (e.g. lift) as follows:
```{r}
redundant <- is.redundant(rules, measure = "lift")
```

```{r}
# remove redundant rules - i.e. exclude rules in "redundant" from set 
rules <- rules[!redundant]
```

```{r}
plot(rules)
```
```{r}
plot(rules, measure = "support", shading = "lift")
```

We could also use the shading to show the number of items in the itemsets, i.e. the “order” of the rule.
```{r}
plot(rules, shading="order")
```
To include a title for you plot you can use main in the control list as follows
```{r}
plot(rules, shading="order", 
control=list(main = "Rule size - Support vs confidence"))
```





```{r}
plot(rules, measure = c("support", "lift"), shading = "order",
     control = list(main = "Support vs Lift, Shaded by Itemset Size"))
```

Plotting an interactive tool
```{r}
plot(rules, measure=c("support", "lift"), 
shading="confidence", engine ="interactive")
```
```{r}
confrules <- rules[quality(rules)$confidence >= 0.75]
plot(confrules)
```

4.	Plot rules with a minimum confidence of 70% and a minimum lift of 6.
```{r}
# Generate rules with specific confidence and lift parameters
rules <- apriori(Groceries, parameter=list(support=0.001, confidence=0.7))  # minimum confidence of 70%

# Filter rules with a minimum lift of 6
filtered_rules <- subset(rules, lift >= 6)

# Plot the rules with specific measurements
plot(filtered_rules, measure = c("support", "lift"), shading = "confidence",
     control = list(main = "Rules with min Confidence of 70% and min Lift of 6"))
```

5.	It is often the case that the most interesting rules are in the support/confidence border.  Obtain an interactive support vs. confidence plot and inspect the border. Are there any interesting rules?
```{r}
plot(rules, measure=c("support", "confidence"), 
shading="confidence", engine ="interactive")
```

```{r}
confrules <- rules[quality(rules)$confidence >= 0.8]
```

```{r}
plot(confrules, method="matrix", measure="lift")
```

```{r}
plot(confrules, method="matrix", measure="lift", 
		control=list(reorder='support/confidence'))
```

6. Plot a matrix of rules with a lift of at least 7. Use confidence as your measure.
```{r}
confrules <- rules[quality(rules)$lift <= 6]
```

```{r}
plot(confrules, method="matrix", measure="confidence", 
		control=list(reorder='support/confidence'))
```

3D plot
```{r}
plot(confrules, method="matrix3D", measure="lift")
```

```{r}
plot(confrules, method="matrix3D", measure="lift", 
		control=list(reorder='support/confidence'))
```
You can plot 2 measures together, one using colour and the other using luminance. Try.
```{r}
plot(confrules, method="matrix", measure=c("lift", "confidence"))
```

and reordering for changed visuals
```{r}
plot(confrules, method="matrix", measure=c("lift", "confidence"), control=list(reorder='similarity'))
```
Plotting best 10 rules
```{r}
liftrules <- head(sort(rules, by="lift"), 10)
plot(liftrules, method="paracoord")
```

```{r}
plot(liftrules, method="paracoord", control=list(reorder=TRUE))
```
7.	Select the best 100 rules  according to support
```{r}
liftrules <- head(sort(rules, by="support"), 100)
plot(liftrules, method="paracoord")
```
8.	Keep only those with a confidence of 60% and a lift of 3.
```{r}
rules <- apriori(Groceries, parameter=list(support=0.001, confidence=0.60))  # minimum confidence of 60%
filtered_rules <- subset(rules, lift >=3 )

# Plot the rules with specific measurements
plot(filtered_rules, measure = "support", shading = "confidence",
     control = list(main = "Rules with min Confidence of 60% and min Lift of 3"))
```

```{r}
butterWhiteBread <- apriori(Groceries, 
parameter=list(support=0.001, confidence=0.5),
	appearance = list(rhs=c("butter", "white bread"), default="lhs"),
	control = list(verbose=F))

inspect(head(sort(butterWhiteBread, by ="lift"),15))

```

```{r}
butterWhiteBreadWholeMilk <- apriori(Groceries, 
parameter=list(support=0.001, confidence=0.5),
appearance = list(lhs=c("butter", "white bread"), 
rhs=c("whole milk")),
			control = list(verbose=F))

inspect(head(sort(butterWhiteBreadWholeMilk, by ="lift"),15))


```

