---
title: "Sample 2"
author: "MS"
date: ""
output: html_document
---
  
  ```{r  setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Clearing the environment.

```{r}
rm(list=ls())
```

Loading  library (package)  tidyverse plus other libraries

```{r warning = F, message = F}

library(tidyverse)
library(caret)
#library(arules)
library(arulesViz)
```

### Loading the data files.


```{r}
contact <- read.csv("contactLenses.csv", header=T, stringsAsFactors=T)
```

```{r}
carprice <- read.csv("CarPrice.csv", header=T, stringsAsFactors=T)
```

```{r}
View(carprice)
```

# Task1
## Task 1a)
### Dataset with 35% data. Use class as the class attribute

```{r}
set.seed(123)
carprice35 <- createDataPartition(y = carprice$carbody, p = .35, list = FALSE)
carprice35 <- carprice[carprice35,]
write.csv(carprice35, file="carprice35.csv", row.names = F)
```




### Define control method using 3 repeats of 10-fold cross validation
```{r}
# setting the validation control - 3 repeats of 10-fold cross validation

ctrl <- trainControl(method = "repeatedcv", repeats = 3, 
                     number = 12,  verboseIter=F)
```


## RF with diabetes dataset

```{r}
# running RF on the carprice dataset
set.seed(123)
my.rf <- train(carbody ~ ., data = carprice,
             method = "rf",
             trControl = ctrl)

# to check accuracy
print(my.rf$results)
```


## NB with the reduced size dataset

```{r}
set.seed(123)
my.rf35 <- train(carbody ~ ., data = carprice35,
             method = "rf",
             trControl = ctrl)

# to check accuracy
print(my.rf35$results)	
# to check size of tree
print(my.rf35$finalModel)

#Confusion matrix
confusionMatrix.train(my.rf35)
```


## kNN with diabetes dataset

```{r}
set.seed(123)
my.knn <- train(carbody ~ ., data = carprice,
             method = "knn",
		 tuneGrid=expand.grid(.k=1:7),preProcess= c("center","scale"),
             trControl = ctrl)
			 
# to check accuracy
print(my.knn$results)	

#Confusion matrix
confusionMatrix(my.knn)
```

## kNN with the reduced size dataset

```{r}
set.seed(123)
my.knn35 <- train(carbody ~ ., data = carprice35,
             method = "knn", preProcess=c("center","scale"),
		 tuneGrid=expand.grid(.k=1:7),
             trControl = ctrl)
			 
# to check accuracy
print(my.knn35$results)	

#Confusion matrix
confusionMatrix(my.knn35)
```
## Comparison of results

```{r}
# comparing all results (95% confidence interval)
results <- resamples(list(Randomforest=my.rf, kNNmodel=my.knn, Randomforest35=my.rf35,kNNmodel35=my.knn35))
scales <- list(x=list(relation="free"), y=list(relation= "free"))
#dotplot(results, scales=scales, conf.level = 0.80)

# comparing all results (80% confidence interval)
dotplot(results, scales=scales, conf.level = 0.80)
```			 
 
### explain the observations in detail based on the results which you were received.
 

 #Task2
 
```{r}
# run the apriori algorithm in the user dataset
set.seed(123)
rules = apriori(contact, parameter=list(support=0.2, confidence=0.95));
```
## Task 2a)
```{r}
inspect(head(sort(rules, by ="confidence"),4))
```
 ## Task 2b)
```{r}
# run the apriori algorithm in the user dataset
set.seed(123)
rules1 = apriori(contact, parameter=list(support=0.1, confidence=0.6));
```
## Task 2a)
```{r}
inspect(head(sort(rules1, by ="lift"),30))
```

## 
```{r}
# sort rules by lift
set.seed(123)
rules.sorted <- sort(rules1, by="lift")


# remove redundant rules

redundant <- is.redundant(rules.sorted, measure = "lift")
rules.pruned <- rules.sorted[!redundant]
rules.pruned
```

## Task 2b) 1. is 
```{r}
# occupation on the RHS
irules <- apriori(contact, parameter=list(support=0.1, confidence=0.6),
appearance = list(rhs=c("tearProductionRate=normal","tearProductionRate=reduced"), default="lhs"))
inspect(head(sort(irules, by ="lift"),10))
```


```{r}
data("Groceries")
summary(Groceries)
View(Groceries)
```
```{r}
rules <- apriori(Groceries, parameter=list(support=0.001, confidence=0.5))
```
```{r}
inspect(head(sort(rules, by ="lift"),10))
```

```{r}
butterWhiteBreadWholeMilk <- apriori(Groceries, 
parameter=list(support=0.001, confidence=0.5),
appearance = list(lhs=c("butter", "white bread"), 
rhs=c("whole milk")),
			control = list(verbose=F))

```
```{r}
inspect(head(sort(butterWhiteBreadWholeMilk, by ="lift"),15))
```

